Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-05-25 paper/paper.Rmd
minor edits; added "temporary" references

This cross-validation bias is well-known in spatial as well as non-spatial prediction [@brenning2005; @brenning2008; @arlot2010; @roberts2017].
It is most easily understood from a predictive modelling perspective by focusing on the question where (and when) the model should be used for prediction.
In crop classification from remotely-sensed data, for instance, learning samples routinely contain multiple grid cells from a sample of fields with known crop type, for instance 2000 grid cells from 100 fields scattered across a large study region.
The purpose of training a model on this sample is to make predictions on other, new *fields* within the same geographic domain (*intra-domain* prediction; @brenning2005) --- not *within* the same field, which obviously presents only one crop type.
In this stituation it would therefore seem rather foolish to train a model on a simple random subsample of grid cells, and test it on the remaining data, i.e. using other grid cells from the same fields, as if we wanted to predict within a field. The results from this performance assessment would be over-optimitic, and perhaps badly so.
To mimic the predictive situation for which the model is trained, we'd rather have to resample at the level of fields, not grid cells [@pena2015].
If the model was to be applied to adjacent agricultural districts, i.e. outside the learning sample's spatial domain (*extra-domain* prediction; @brenning2005), it would even seem necessary to resample at a higher level of spatial aggregation, i.e. at the level of agricultural districts within the learning sample, in order to realistically mimic the actual prediction task.
The cross-validation resampling needed therefore depends as much on the prediction task as on the data structure or dependency at hand.

@@

In cross-validation for predictive model assessment, we consider the following formal setting.
We are interested in predicting a numerical or categorical response $y$ of an object or instance using a feature vector $\mathbf{x} = (x^{(1)}, \ldots, x^{(p)})\in\mathbb{R}^p$ and a model $\hat{f}_\mathcal{L}$ that has been trained on a learning sample $\mathcal{L} = \{(y_i, \mathbf{x}_i),\ i = 1, ..., n\}$. Our goal is to estimate the expected value of the performance of $\hat{f}_\mathcal{L}$,
\[
\mathit{perf(\hat{f}_\mathcal{L})} := E(l(Y,\hat{f}_\mathcal{L}(X))),
\]
where $l$ is a real-valued loss function, and the expected value is with respect the probability distribution of $X$, the features of an instance $(Y,X)$ drawn randomly from the underlying population.
This is referred to as the *actual* or *conditional* performance measure as it is conditional on $L$ [@milligan2000].
The loss function can take a variety of forms such as the misclassification error $I(Y\neq\hat{f}_\mathcal{L}(X))$ in classification, or the squared error $(Y-\hat{f}_\mathcal{L}(X))^2$ in regression, among many other possible measures.
The choice of the performance measure is equally critical as the choice of the estimation procedure, but it is beyond the scope of this contribution to discuss performance metrics for regression and classification (see e.g., @hand1997 for classification, and ref?? for regression tasks).

>>>>

This cross-validation bias is well-known in spatial as well as non-spatial prediction [@brenning2005; @brenning2008; @arlot2010; @roberts2017] and in forecasting (ref FIXME: <https://doi.org/10.1016/j.csda.2017.11.003>).
It is most easily understood from a predictive modelling perspective by focusing on the question where (and when) the model should be used for prediction.
In crop classification from remotely-sensed data, for instance, learning samples routinely contain multiple grid cells from a sample of fields with known crop type, for instance 2000 grid cells from 100 fields scattered across a large study region.
The purpose of training a model on this sample is to make predictions on other, new *fields* within the same geographic domain (*intra-domain* prediction; @brenning2005) --- not *within* the same field, which obviously presents only one crop type.
In this stituation it would therefore seem rather foolish to train a model on a simple random subsample of grid cells, and test it on the remaining data, i.e. using other grid cells from the same fields, as if we wanted to predict within a field. The results from this performance assessment would be over-optimitic, and perhaps badly so.
To mimic the predictive situation for which the model is trained, we'd rather have to resample at the level of fields, not grid cells [@pena2015].
If the model was to be applied to adjacent agricultural districts, i.e. outside the learning sample's spatial domain (*extra-domain* prediction; @brenning2005), it would even seem necessary to resample at a higher level of spatial aggregation, i.e. at the level of agricultural districts within the learning sample, in order to realistically mimic the actual prediction task.
The cross-validation resampling needed therefore depends as much on the prediction task as on the data structure or dependency at hand.

@@

In cross-validation for predictive model assessment, we consider the following formal setting.
We are interested in predicting a numerical or categorical response $y$ of an object or instance using a feature vector $\mathbf{x} = (x^{(1)}, \ldots, x^{(p)})\in\mathbb{R}^p$ and a model $\hat{f}_\mathcal{L}$ that has been trained on a learning sample $\mathcal{L} = \{(y_i, \mathbf{x}_i),\ i = 1, ..., n\}$. Our goal is to estimate the expected value of the performance of $\hat{f}_\mathcal{L}$,
\[
\mathit{perf(\hat{f}_\mathcal{L})} := E(l(Y,\hat{f}_\mathcal{L}(X))),
\]
where $l$ is a real-valued loss function, and the expected value is with respect the probability distribution of $X$, the features of an instance $(Y,X)$ drawn randomly from the underlying population.
This is referred to as the *actual* or *conditional* performance measure as it is conditional on $L$ [@milligan2000].
The loss function can take a variety of forms such as the misclassification error $I(Y\neq\hat{f}_\mathcal{L}(X))$ in classification, or the squared error $(Y-\hat{f}_\mathcal{L}(X))^2$ in regression, among many other possible measures.
The choice of the performance measure is equally critical as the choice of the estimation procedure, but it is beyond the scope of this contribution to discuss performance metrics for regression and classification (see e.g., @hand1997 for classification, and ref FIXME perhaps <https://doi.org/10.1016/j.ijforecast.2006.03.001> for regression and forecasting tasks).
